Effects of Different Pain Relievers and Dosages on Cognitive Retention

### 2025-03-22 paper/paper.qmd
final changes

## Research Questions
1. *How does the type of pain reliever (Aspirin 500 mg, Paracetamol 500 mg, Tramadol 50 mg, and Placebo) affect cognitive task performance?*
- Null Hypothesis: All pain relievers at normal dosages act the same for cognitive retention game performances.
- Alternative Hypothesis: There exists at least one pair of distinct pain relievers at normal dosages which act differently for cognitive retention game performances.

2. *How does the dosage (low vs. high) of each drug affect cognitive retention, while accounting for confounding factors like age?*
- Null Hypothesis: There is no interaction effect between the type of pain reliever and dosage level on the improvement in memory task performance.
- Alternative Hypothesis: There is an interaction effect between at least one of the types of pain reliever and dosage level on the improvement in memory task performance.

## Study Design  
This study involves 2 factors, pain reliever type (4 levels) and dosage(2 levels). Crossing them provides us with *8 treatment groups*, with **30 participants** assigned to each. The groups are as follows:

@@

## Confounding Variable: Age  
To reduce bias, we consider to control the experiment by controlling participant ages to be 18+. Hence in this study **age** acts as a **confounding variable**.
Participants will be categorized into three groups evenly to prevent bias created through age. 
The groups are as follows:

1. 18–34 years
2. 35–50 years
3. 50+ years

@@

## Data Analysis Method   
We will use **one-way ANOVA** to analyze the effects of different pain relievers on memory performance.
Similarly, we will use **two-way ANOVA** to analyze the effects of different dosage levels per drug type on memory performance and cognitive retention.

@@

# Summary Statistics

>>>>

## Research Questions
1. *How does the type of pain reliever (Aspirin 500 mg, Paracetamol 500 mg, Tramadol 50 mg, and Placebo) affect cognitive retention game performance?*
- Null Hypothesis: All pain relievers at normal dosages act the same for cognitive retention game performances.
- Alternative Hypothesis: There exists at least one pair of distinct pain relievers at normal dosages which act differently for cognitive retention game performances.

2. *Is there an interaction effect between the type of pain reliever and dosage level on the improvement in memory task performance?*
- Null Hypothesis: There is no interaction effect between the type of pain reliever and dosage level on the improvement in memory task performance.
- Alternative Hypothesis: There is an interaction effect between at least one of the types of pain reliever and dosage level on the improvement in memory task performance.

# Methodology

In February, our group members collected observations from participants belonging to a virtual population to assess whether the treatment had any effect on memory performance, this resulted in a total of 240 observations. Each participant was **randomly** assigned a treatment.

The experiment utilized a two-way ANOVA design, where the experimental unit was the participant, the explanatory variables were the type of painkiller and dosage level (low/high), and the response variable was the memory performance, measured by the memory game score.

## Study Design and Treatments   
We selected three brands of painkillers: Aspirin, Paracetamol, and Tramadol. Additionally, we considered a placebo treatment. As a result, this study involves 2 factors: pain reliever type (4 levels - Aspirin, Paracetamol, Tramadol, Placebo) and dosage (2 levels - Low, High). Crossing the two factors provides us with a total of *8 treatment groups*, with **30 participants** assigned to each. The groups are as follows:

@@

For each treatment, we selected 30 different people (n = 30). Through randomly picking the virtual citizens, we replicated 30 experimental units per treatment. The even observations per treatment group ensures we have a balanced design.

### Nuisance Variable: Age  
To reduce bias, we consider **controlling** the experiment by limiting participant ages to be strictly 18+. Hence in this study **age** acts as a **nuisance variable**.
We also used **blocking** for age by ensuring participants were categorized into three even groups to prevent bias created through age. 
The groups are as follows:
1. 18–34 years old.
2. 35–50 years old.
3. 50+ years old.

@@

@@
